### 2023-06-22 tex/trans.tex
[ writing ] renew generic syntax chapter
[ edit ] inferrule -> ebproof

\subsection{Dual Intuitionistic Linear Logic}\label{sec:dill}

@@

The types of DILL are the same as the types of \name, except for the
restriction of $\oc\gr{r}$ to just $\oc\gr\omega$.
I will write the latter simply as $\oc$ when it appears in DILL\@.
I add sums and with-products to the calculus of \cite{Barber1996}, with the
obvious rules (stated fully in \cref{fig:dill}).
These additive type formers present no additional difficulty to the translation.

\begin{figure}
  \begin{mathpar}
    \inferrule*[right=Int-Ax]{ }
    {\gamma, A; \cdot \vdash A}
    \and
    \inferrule*[right=Lin-Ax]{ }
    {\gamma; A \vdash A}
    \and
    \inferrule*[right=$I$-I]{ }
    {\gamma; \cdot \vdash I}
    \and
    \inferrule*[right=$I$-E]
    {\gamma; \delta_1 \vdash I \\ \gamma; \delta_2 \vdash A}
    {\gamma; \delta_1, \delta_2 \vdash A}
    \and
    \inferrule*[right=$\otimes$-I]
    {\gamma; \delta_1 \vdash A \\ \gamma, \delta_2 \vdash B}
    {\gamma; \delta_1, \delta_2 \vdash A \otimes B}
    \and
    \inferrule*[right=$\otimes$-E]
    {\gamma; \delta_1 \vdash A \otimes B \\ \gamma; \delta_2, A, B \vdash C}
    {\gamma; \delta_1, \delta_2 \vdash C}
    \and
    \inferrule*[right=$\multimap$-I]
    {\gamma; \delta, A \vdash B}
    {\gamma; \delta \vdash A \multimap B}
    \and
    \inferrule*[right=$\multimap$-E]
    {\gamma; \delta_1 \vdash A \multimap B \\ \gamma; \delta_2 \vdash A}
    {\gamma; \delta_1, \delta_2 \vdash B}
    \and
    \inferrule*[right=$\oc$-I]
    {\gamma; \cdot \vdash A}
    {\gamma; \cdot \vdash \oc A}
    \and
    \inferrule*[right=$\oc$-E]
    {\gamma; \delta_1 \vdash \oc A \\ \gamma, A; \delta_2 \vdash B}
    {\gamma; \delta_1, \delta_2 \vdash B}
    \and
    \inferrule*[right=$\top$-I]{ }
    {\gamma; \delta \vdash \top}
    \and
    \inferrule*[right=$\with$-I]
    {\gamma; \delta \vdash A \\ \gamma, \delta \vdash B}
    {\gamma; \delta \vdash A \with B}
    \and
    \inferrule*[right=$\with$-E$_i$]
    {\gamma; \delta \vdash A_0 \with A_1}
    {\gamma; \delta \vdash A_i}
    \and
    \inferrule*[right=$0$-E]
    {\gamma; \delta_1 \vdash 0}
    {\gamma; \delta_1, \delta_2 \vdash A}
    \and
    \inferrule*[right=$\oplus$-I$_i$]
    {\gamma; \delta \vdash A_i}
    {\gamma; \delta \vdash A_0 \oplus A_1}
    \and
    \inferrule*[right=$\oplus$-E]
    {
      \gamma; \delta_1 \vdash A \oplus B \\
      \gamma; \delta_2, A \vdash C \\
      \gamma; \delta_2, B \vdash C
    }
    {\gamma; \delta_1, \delta_2 \vdash C}
  \end{mathpar}
  \label{fig:dill}
  \caption{The rules of DILL, extended with additive connectives}
\end{figure}

@@

  Weakening is used when splitting linear variables between two premises.
  For example, \TirName{$\otimes$-I} in DILL is as follows.
  \[
    \inferrule*[right=$\otimes$-I]
    {\gamma; \delta_t \vdash t : A \\ \gamma; \delta_u \vdash u : B}
    {\gamma; \delta_t, \delta_u \vdash t \otimes u : A \otimes B}
  \]
  From this, our new derivation is as follows.
  \[
    \inferrule*[right=$\otimes$-I]
    {
      \inferrule*[right=Weak]
      {\mathit{ih}_t \\\\
        \gr\omega\gamma, \gr1\delta_t
        \vdash M_t : A}
      {\gr\omega\gamma, \gr1\delta_t,
        \gr0\delta_u
        \vdash M_t : A}
      \\
      \inferrule*[right=Weak]
      {\mathit{ih}_u \\\\
        \gr\omega\gamma, \gr1\delta_u
        \vdash M_u : A}
      {\gr\omega\gamma, \gr0\delta_t,
        \gr1\delta_u
        \vdash M_u : A}
    }
    {\gr\omega\gamma, \gr1\delta_t,
      \gr1\delta_u
      \vdash \plr{M_t, M_u} : A \otimes B}
  \]
\end{proof}

@@

\[
  \inferrule*[right=$\otimes$-I]
  {
    \inferrule*[right=var]{ }
    {\gr1A : x \vdash A}
    \\
    \inferrule*[right=var]{ }
    {\gr1A : x \vdash A}
    \\
    \gr\omega \leq \gr1 + \gr1
  }
  {\gr\omega A : x \vdash A \otimes A}
\]

@@

\begin{definition}
  A derivation is said to be \emph{$\instDILL$-bottom-up} if only the following
  facts about addition and multiplication are used, and all proofs of
  inequalities not at leaves are by reflexivity (i.e, not using the facts that
  $\gr\omega \leq \gr0$ and $\gr\omega \leq \gr1$).

@@

\subsection{Pfenning Davies}\label{sec:pd}

@@

\begin{figure}
  \begin{mathpar}
    \inferrule*[right=hyp]{ }
    {\gamma; \delta, A\;\mathit{true} \vdash A\;\mathit{true}}
    \and
    \inferrule*[right=hyp*]{ }
    {\gamma, A\;\mathit{valid}; \delta \vdash A\;\mathit{true}}
    \and
    \inferrule*[right=$\supset$I]
    {\gamma; \delta, A\;\mathit{true} \vdash B\;\mathit{true}}
    {\gamma; \delta \vdash A \supset B\;\mathit{true}}
    \and
    \inferrule*[right=$\supset$E]
    {
      \gamma; \delta \vdash A \supset B\;\mathit{true} \\
      \gamma; \delta \vdash A\;\mathit{true}
    }
    {\gamma; \delta \vdash B\;\mathit{true}}
    \and
    \inferrule*[right=$\Box$I]
    {\gamma; \cdot \vdash A\;\mathit{true}}
    {\gamma; \delta \vdash \Box A\;\mathit{true}}
    \and
    \inferrule*[right=$\Box$-E]
    {
      \gamma; \delta \vdash \Box A\;\mathit{true} \\
      \gamma, A\;\mathit{valid}; \delta \vdash B\;\mathit{true}
    }
    {\gamma; \delta \vdash B\;\mathit{true}}
    \and
    \inferrule*[right=$\top$-I]{ }
    {\gamma; \delta \vdash \top\;\mathit{true}}
    \and
    \inferrule*[right=$\wedge$-I]
    {
      \gamma; \delta \vdash A\;\mathit{true} \\
      \gamma, \delta \vdash B\;\mathit{true}
    }
    {\gamma; \delta \vdash A \wedge B\;\mathit{true}}
    \and
    \inferrule*[right=$\wedge$-E$_i$]
    {\gamma; \delta \vdash A_0 \wedge A_1\;\mathit{true}}
    {\gamma; \delta \vdash A_i\;\mathit{true}}
    \and
    \inferrule*[right=$\bot$-E]
    {\gamma; \delta \vdash \bot\;\mathit{true}}
    {\gamma; \delta \vdash A\;\mathit{true}}
    \and
    \inferrule*[right=$\vee$-I$_i$]
    {\gamma; \delta \vdash A_i\;\mathit{true}}
    {\gamma; \delta \vdash A_0 \vee A_1\;\mathit{true}}
    \and
    \inferrule*[right=$\vee$-E]
    {
      \gamma; \delta \vdash A \vee B\;\mathit{true} \\
      \gamma; \delta, A \vdash C\;\mathit{true} \\
      \gamma; \delta, B \vdash C\;\mathit{true}
    }
    {\gamma; \delta \vdash C\;\mathit{true}}
  \end{mathpar}
  \label{fig:pd}
  \caption{The rules of PD, extended with several standard connectives}
\end{figure}

>>>>

\subsection{Dual Intuitionistic Linear Logic}\label{sec:trans-dill}

@@

The types of DILL are the same as the types of \name, except for the
restriction of $\oc\gr{r}$ to just $\oc\gr\omega$.
I will write the latter simply as $\oc$ when it appears in DILL\@.
I add sums and with-products to the calculus of \cite{Barber1996}, with the
obvious rules (stated fully in \cref{fig:dill2}).
These additive type formers present no additional difficulty to the translation.

\begin{figure}
  \begin{mathpar}
    \ebrule{\infer0[Int-Ax]{\gamma, A; {\cdot} \vdash A}} \and
    \ebrule{\infer0[Lin-Ax]{\gamma; A \vdash A}} \and
    \ebrule{\infer0[$I$-I]{\gamma; {\cdot} \vdash I}} \and
    \ebrule{%
      \hypo{\gamma; \delta_1 \vdash I}
      \hypo{\gamma; \delta_2 \vdash A}
      \infer2[$I$-E]{\gamma; \delta_1, \delta_2 \vdash A}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta_1 \vdash A}
      \hypo{\gamma, \delta_2 \vdash B}
      \infer2[$\otimes$-I]{\gamma; \delta_1, \delta_2 \vdash A \otimes B}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta_1 \vdash A \otimes B}
      \hypo{\gamma; \delta_2, A, B \vdash C}
      \infer2[$\otimes$-E]{\gamma; \delta_1, \delta_2 \vdash C}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta, A \vdash B}
      \infer1[$\multimap$-I]{\gamma; \delta \vdash A \multimap B}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta_1 \vdash A \multimap B}
      \hypo{\gamma; \delta_2 \vdash A}
      \infer2[$\multimap$-E]{\gamma; \delta_1, \delta_2 \vdash B}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \cdot \vdash A}
      \infer1[$\oc$-I]{\gamma; \cdot \vdash \oc A}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta_1 \vdash \oc A}
      \hypo{\gamma, A; \delta_2 \vdash B}
      \infer2[$\oc$-E]{\gamma; \delta_1, \delta_2 \vdash B}
    }
    \and
    \ebrule{%
      \infer0[$\top$-I]{\gamma; \delta \vdash \top}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta \vdash A}
      \hypo{\gamma, \delta \vdash B}
      \infer2[$\with$-I]{\gamma; \delta \vdash A \with B}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta \vdash A_0 \with A_1}
      \infer1[$\with$-E$_i$]{\gamma; \delta \vdash A_i}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta_1 \vdash 0}
      \infer1[$0$-E]{\gamma; \delta_1, \delta_2 \vdash A}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta \vdash A_i}
      \infer1[$\oplus$-I$_i$]{\gamma; \delta \vdash A_0 \oplus A_1}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta_1 \vdash A \oplus B}
      \hypo{\gamma; \delta_2, A \vdash C}
      \hypo{\gamma; \delta_2, B \vdash C}
      \infer3[$\oplus$-E]{\gamma; \delta_1, \delta_2 \vdash C}
    }
  \end{mathpar}
  \label{fig:dill2}
  \caption{The rules of DILL, extended with additive connectives}
\end{figure}

@@

  Weakening is used when splitting linear variables between two premises.
  For example, \TirName{$\otimes$-I} in DILL is as follows.
  \[
    \ebrule{%
      \hypo{\gamma; \delta_t \vdash t : A}
      \hypo{\gamma; \delta_u \vdash u : B}
      \infer2[$\otimes$-I]%
      {\gamma; \delta_t, \delta_u \vdash t \otimes u : A \otimes B}
    }
  \]
  From this, our new derivation is as follows.
  \[
    \ebrule{%
      \hypo{\mathit{ih}_t}
      \infer[no rule]1{\gr\omega\gamma, \gr1\delta_t \vdash M_t : A}
      \infer1[Weak]%
      {\gr\omega\gamma, \gr1\delta_t, \gr0\delta_u \vdash M_t : A}
      \hypo{\mathit{ih}_u}
      \infer[no rule]1{\gr\omega\gamma, \gr1\delta_u \vdash M_u : A}
      \infer1[Weak]%
      {\gr\omega\gamma, \gr0\delta_t, \gr1\delta_u \vdash M_u : A}
      \infer2[$\otimes$-I]%
      {\gr\omega\gamma, \gr1\delta_t, \gr1\delta_u
        \vdash \plr{M_t, M_u} : A \otimes B}
    }
  \]
\end{proof}

@@

\[
  \ebrule{%
    \infer0[Var]{\gr1A : x \vdash A}
    \infer0[Var]{\gr1A : x \vdash A}
    \hypo{\gr\omega \leq \gr1 + \gr1}
    \infer3[$\otimes$-I]{\gr\omega A : x \vdash A \otimes A}
  }
\]

@@

\begin{definition}\label{def:DILL-bottom-up}
  A derivation is said to be \emph{$\instDILL$-bottom-up} if only the following
  facts about addition and multiplication are used, and all proofs of
  inequalities not at leaves are by reflexivity (i.e, not using the facts that
  $\gr\omega \leq \gr0$ and $\gr\omega \leq \gr1$).

@@

\subsection{Pfenning Davies}\label{sec:trans-pd}

@@

\begin{figure}
  \begin{mathpar}
    \ebrule{%
      \infer0[hyp]{\gamma; \delta, A\;\mathit{true} \vdash A\;\mathit{true}}
    }
    \and
    \ebrule{%
      \infer0[hyp*]{\gamma, A\;\mathit{valid}; \delta \vdash A\;\mathit{true}}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta, A\;\mathit{true} \vdash B\;\mathit{true}}
      \infer1[$\supset$I]{\gamma; \delta \vdash A \supset B\;\mathit{true}}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta \vdash A \supset B\;\mathit{true}}
      \hypo{\gamma; \delta \vdash A\;\mathit{true}}
      \infer2[$\supset$E]{\gamma; \delta \vdash B\;\mathit{true}}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \cdot \vdash A\;\mathit{true}}
      \infer1[$\Box$I]{\gamma; \delta \vdash \Box A\;\mathit{true}}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta \vdash \Box A\;\mathit{true}}
      \hypo{\gamma, A\;\mathit{valid}; \delta \vdash B\;\mathit{true}}
      \infer2[$\Box$-E]{\gamma; \delta \vdash B\;\mathit{true}}
    }
    \and
    \ebrule{%
      \infer0[$\top$-I]{\gamma; \delta \vdash \top\;\mathit{true}}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta \vdash A\;\mathit{true}}
      \hypo{\gamma, \delta \vdash B\;\mathit{true}}
      \infer2[$\wedge$-I]{\gamma; \delta \vdash A \wedge B\;\mathit{true}}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta \vdash A_0 \wedge A_1\;\mathit{true}}
      \infer1[$\wedge$-E$_i$]{\gamma; \delta \vdash A_i\;\mathit{true}}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta \vdash \bot\;\mathit{true}}
      \infer1[$\bot$-E]{\gamma; \delta \vdash A\;\mathit{true}}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta \vdash A_i\;\mathit{true}}
      \infer1[$\vee$-I$_i$]{\gamma; \delta \vdash A_0 \vee A_1\;\mathit{true}}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta \vdash A \vee B\;\mathit{true}}
      \hypo{\gamma; \delta, A \vdash C\;\mathit{true}}
      \hypo{\gamma; \delta, B \vdash C\;\mathit{true}}
      \infer3[$\vee$-E]{\gamma; \delta \vdash C\;\mathit{true}}
    }
  \end{mathpar}
  \label{fig:pd}
  \caption{The rules of PD, extended with several standard connectives}
\end{figure}
